Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-06-21 manuscript/manuscript_06-20-19.tex
section 3

We note that the multivariate skew normal density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005). As we will see in Section 3.6, this specification facilitates prior specification by admitting convenient joint prior distributions for the regression coefficients and scale matrices. In particular, we define $\mathbf{Y}_k$ as the ${n_k \times J}$  response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$). 
$$\mathbf{Y}_{k} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)$$
$$\text{vec}(\mathbf{Y}) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
$$\text{vec}(\mathbf{M}) = (\boldsymbol\zeta_1^T,...,\boldsymbol\zeta_{n_k}^T)$$
where $\boldsymbol\mu_i = \mathbf{x}_i^T \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$. This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\zeta_i, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$. Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$. This implies that all responses for cluster $k$ conditioned on $\mathbf{t}_{k}$ is distributed as 
$$\mathbf{Y}_{n_k \times J} | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),$$
where $\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ and $\mathbf{X}_k$ is an $n_k \times p$ design matrix.

>>>>

We note that the multivariate skew normal density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005). As we will see in Section 3.6, this specification facilitates prior specification by admitting convenient joint prior distributions for the regression coefficients and scale matrices. In particular, we define $\mathbf{Y}_k$ as the ${n_k \times J}$  response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$). 
$$\mathbf{Y}_{k} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)$$
$$\text{vec}(\mathbf{Y}) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
$$\text{vec}(\mathbf{M}) = (\boldsymbol\zeta_1^T,...,\boldsymbol\zeta_{n_k}^T)$$
where $\boldsymbol\zeta_i = \mathbf{x}_i^T \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$. This implies that the $i^{th}$ row of $\mathbf{Y}_k$, a ${n_k \times J}$ matrix of responses for cluster $k$, is $\mathbf{y}_i \sim MSN_J(\boldsymbol\zeta_i, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$. Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$. This implies that all responses for cluster $k$ conditioned on $\mathbf{t}_{k}$ is distributed as 
$$\mathbf{Y}_k | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),$$
where $\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ and $\mathbf{X}_k$ is an $n_k \times p$ design matrix.
